---
title: "Lab 3 Data Manipulation using dplyr package"
author: "Makesh Srinivasan"
date: "8/17/2021"
output: html_document
---
Registration number: 19BCE1717 <br>
Faculty: Dr. C. Sweetlin Hemalatha <br>
Slot: L39 + L40 <br>
Course code: CSE3505 <br>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Obtain a data frame called “newsurvey” by removing all ‘na’ values in “survey” data of MASS package to do the following:
```{r}
library(MASS)
data(survey)
print("Dimensions before omitting na values")
dim(survey)

newsurvey = na.omit(survey)
print("Dimensions after omitting na values")
dim(newsurvey)
head(newsurvey)
```

1.	Install the dplyr package and import it.
```{r}
# The following comment contains the code to install the package
# install.packages("dplyr")
library(dplyr)
# Installed the package
```

2.	Filter the top 5 male left handers.	
```{r}
head(filter(newsurvey, Sex == "Male", W.Hnd == "Left"),5)
```

3.	Display all female right handers who keep right on top while clapping.
```{r}
# View function will display all the matching entries but it does not get displayed on HTML page. Hence the View is removed
# View(filter(newsurvey, Sex == "Female", W.Hnd == "Right", Clap == "Right"))
filter(newsurvey, Sex == "Female", W.Hnd == "Right", Clap == "Right")
```

4.	Display all students who never exercise.
```{r}
# View function will display all the matching entries but it does not get displayed on HTML page. Hence the View is removed
# View(filter(newsurvey, Exer == "None"))
filter(newsurvey, Exer == "None")
```

5.	Display random 5 rows of gender, age and writing hand columns of the students.
```{r}
sample_n(newsurvey[c("Sex","Age","W.Hnd")], 5)
```

6.	Display random 50% of the rows of age, pulse rate and writing hand span of female left handers.
```{r}
newsurvey %>%
  filter(Sex == "Female", W.Hnd == "Left") %>%
  dplyr::sample_frac(0.5) %>%
  dplyr::select(Age, Pulse, Wr.Hnd)
```

7.	Arrange all male left handers according to descending order of their heights.
```{r}
dplyr::arrange(filter(newsurvey, Sex == "Male", W.Hnd == "Left"), desc(Height))
```

8.	Introduce a new column hand_span which contains the value as difference between the span of writing hand and non-writing hand and display it along with gender, writing hand and non-writing hand span.
```{r}
newsurvey%>%
  mutate(hand_span=Wr.Hnd-NW.Hnd) %>%
  select(Sex, W.Hnd, NW.Hnd, hand_span)
```

9.	Display the average writing span of male and female left handers.
```{r}
newsurvey %>%
  filter(W.Hnd == "Left") %>%
  group_by(Sex) %>%
  summarize(Average = mean(Wr.Hnd))

# The answer was verified using the following lines of code as well
# print(paste("Average writing span of male left handers = ", mean(filter(newsurvey, Sex=="Male", W.Hnd=="Left")$Wr.Hnd)))
# print(paste("Average writing span of female left handers = ", mean(filter(newsurvey, Sex=="Female", W.Hnd=="Left")$Wr.Hnd)))

```

10.	Find the maximum pulse rate of male left and right handers.
```{r}
newsurvey %>%
  filter(Sex == "Male") %>%
  group_by(W.Hnd) %>%
  summarize(Maximum = max(Pulse))
```


