---
title: "Lab exercise 1: Basic Data Types in R"
author: "Makesh Srinivasan"
date: "8/8/2021"
output: html_document
---
Registration number: 19BCE1717 <br />
Faculty: Dr. C. Sweetlin Hemalatha <br />
Slot: L39 + L40 <br />
Course code: CSE3505 <br />

## Solve the following questions using R
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Q1) Convert TRUE to an integer
```{r}
print("Before conversion: ")
x <- TRUE
class(x)

print("After conversion: ")
x <- as.integer(x)
class(x)
```
Q2) Make this quote into an R string
– “Do you think this is a game?”, he said. “No, I think Jenga’s a game”, Archer responded.
```{r}
quote_str <-"\"Do you think this is a game?\", he said. \"No, I think Jenga\'s a game\", Archer responded."
writeLines(quote_str)
class(quote_str)
```
Q3)	Assign an integer value to two variables p & q
```{r}
p <- as.integer(1)
q <- as.integer(2)

print(paste("Value of p: ", p))
print(paste("Class of p is: ", class(p)))

print(paste("Value of q: ", q))
print(paste("Class of q is: ", class(q)))
```
Q4)	Use R as a calculator to do basic arithmetic operations on p &q and store each result in different variables.
```{r}
addn <- p+q
subn <- p-q
divn <- p/q
muln <- p*q
mods <- p%%q
expo <- p^q
print(paste("Addition = ", addn))
print(paste("Subtraction = ", subn))
print(paste("Division = ", divn))
print(paste("Multiplication = ", muln))
print(paste("Modulus = ", mods))
print(paste("Exponent = ", expo))

```
Q5) Calculate sqrt(p) and perform log2 transformation on the result.
```{r}
result <- sqrt(p)
print(paste("Square Root: ", result))
print(paste("Log2: ", log2(result)))
```
Q6) Calculate the 10-based logarithm of 100 and multiply the result with cosine of π.
```{r}
answer <- log10(100)*cos(pi)
answer
```
Q7) Create a vector ‘x’ using : operator from -2 to 2.
```{r}
x <- -2:2
x
```
Q8) Create a vector ‘y’ using c() with values 1 to 10. What is the dimension of y and length of y.
```{r}
y <- c(1,2,3,4,5,6,7,8,9,10)
y
print("Dimension:")
dim(y)
print(paste("Length:", length(y)))
```
Q9) Create two vectors small and caps with first 5 alphabets in lower case and last 5 alphabets in upper case respectively. Bind it row and columnwise.
```{r}
small <- c("a","b","c","d","e")
caps <- c("Q","W","E","R","T")
rbind(small,caps)
cbind(small,caps)
```
Q10) Create a matrix (M) and find MxMxM
```{r}
M <- matrix(c(1,-2,5,4,8,-1,3,6,7),nrow = 3,ncol = 3)
print("Matrix:")
M
print("After Matrix Multiplication of MxMxM")
M%*%M%*%M
print("Element-wise Multiplication (not matrix multiplication) of M*M*M")
M*M*M
```
Q11) Find the result of elementwise multiplication of M by M.
```{r}
print("Element-wise multiplication of M*M")
M*M
```
Q12) Find the transpose, inverse and determinant of M.
```{r}
print("Transpose:")
t(M)
print("Inverse:")
solve(M)
print("Determinant:")
det(M)
```