---
title: "Lab 13: Environmental analysis"
author: "Makesh Srinivasan"
date: "11/30/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Environmental analysis
Instructions: Choose data set related to environment of your choice and do the possible analysis. Mention the data set in the submission
source: https://www.kaggle.com/ruchi798/global-environmental-indicators > land and agriculture > Agricultural land.csv

For this exercise, I plan to do the following.
1) Data manipulation and cleaning
2) Regression analysis
3) Visualisation and plotting

NOTE: all the previous exercises (8 in total) are incorporated into these four topics.


NOTE: Since at the begining of the class it was said that there is no marking for this lab exercise, I have no provided inferences and conclusions for the plot or the linear regression models. This is because, I wanted to explore the syntax and plot various plots rather than explain the plots. There is an issue with knitting of HTML document becasue of the LabelEncoder function, and thus, I have uploaded the RMD file.

## Data manipulation and cleaning
Import the dataset:
```{r}
data <- read.csv("Agricultural Land.csv")
head(data)
```

Str:
```{r}
str(data)
```

Get col names:
```{r}
data <- subset(data, select = -c(X))
cols = colnames(data)
```

Libraries:
```{r}
library(dplyr)
library(caret)
library(lattice)
library(e1071)
library(ggplot2)
library(tidyr)
library(stringr)
```

Rename columns:
```{r}
names(data)[2] <- 'Agr_area'
names(data)[3] <- 'part_area'
names(data)[4] <- 'part_land'
names(data)[5] <- 'Arable'
names(data)[6] <- 'crops'
names(data)[7] <- 'meadows'
names(data)[8] <- 'irrigated'
```


Trim the white spaces:
```{r}
trim <- function (x) gsub("^\\s+|\\s+$", "", x)
data$Agr_area <- trim(data$Agr_area)
data$part_area <- trim(data$part_area)
data$part_land <- trim(data$part_land)
data$Arable <- trim(data$Arable)
data$crops <- trim(data$crops)
data$meadows <- trim(data$meadows)
data$irrigated <- trim(data$irrigated)
str(data)
```

Remove the white space between the numbers:
```{r}
mClean <- function(strVec){
  pass1 <- strVec %>% 
    str_trim() %>% 
    str_extract("(?x)        # Perl-style whitespace
                ^[\\+\\-]?   # An optional leading +/-
                \\d+         # the integer part
                (\\.\\d+)? # A fractional part
                ") %>% 
    as.numeric()
}

data <- data%>%
  mutate(
    Agr_area = as.numeric(mClean(Agr_area)), 
    part_area = as.numeric(mClean(part_area)), 
    part_land = as.numeric(mClean(part_land)), 
    Arable = as.numeric(mClean(Arable)), 
    crops = as.numeric(mClean(crops)), 
    meadows = as.numeric(mClean(meadows)), 
    irrigated = as.numeric(mClean(irrigated)))
str(data)
```

Replace NA values
```{r}
data <- data %>%
  mutate(Agr_area = replace(Agr_area, is.na(Agr_area), mean(Agr_area, na.rm = TRUE)))

data <- data %>%
  mutate(part_area = replace(part_area, is.na(part_area), mean(part_area, na.rm = TRUE)))

data <- data %>%
  mutate(part_land = replace(part_land, is.na(part_land), mean(part_land, na.rm = TRUE)))

data <- data %>%
  mutate(Arable = replace(Arable, is.na(Arable), mean(Arable, na.rm = TRUE)))

data <- data %>%
  mutate(crops = replace(crops, is.na(crops), mean(crops, na.rm = TRUE)))

data <- data %>%
  mutate(meadows = replace(meadows, is.na(meadows), mean(meadows, na.rm = TRUE)))

data <- data %>%
  mutate(irrigated = replace(irrigated, is.na(irrigated), mean(irrigated, na.rm = TRUE)))

head(data)
```

Encode the countries
```{r}
lab_enc <- LabelEncoder.fit(data$Country)
data['Country'] <- transform(lab_enc, matrix(data$Country))
```

Str:
```{r}
str(data)
```

NA none:
```{r}
print(paste("Number of missing values = ", sum(is.na(data))))
```

Correlation:
```{r}
cols = colnames(data)
for (c in cols) {
  print(paste("Column", c, ": ", cor(data[[c]],data$Agr_area)))
}
```

Covariance:
```{r}
cols = colnames(data)
for (c in cols) {
  print(paste("Column", c, ": ", cov(data[[c]],data$Agr_area)))
}
```

## Regression analysis

Linear regression model:
```{r}
model1 <- lm(Agr_area~.,data=data)
model1
```
Summary:
```{r}
summary(model1)
```

Sigma and confidence interval:
```{r}
sigma(model1)
confint(model1)
```

Residual vs Fitted
```{r}
plot(model1, 1)
```

```{r}
par(mar=c(2,2,2,2))
hist(model1$residuals)
```

```{r}
qqnorm(model1$residuals,ylab = "Residuals")
qqline(model1$residuals)
```

Statistically significant attributes:
```{r}
sig_data = subset(data, select = c(Arable, meadows, Agr_area))
str(sig_data)
```

Plots:
```{r}
ggplot(data,aes(x=Arable,y=Agr_area))+geom_point()+geom_smooth()
ggplot(data,aes(x=Arable,y=Agr_area))+geom_point()+geom_smooth(method='lm',se=FALSE)
ggplot(data,aes(x=meadows,y=Agr_area))+geom_point()+geom_smooth()
```

train model:
```{r}
set.seed(123)
train_samples <- sig_data$Agr_area %>% 
  createDataPartition(p=0.8,list=FALSE)
head(train_samples)
```

```{r}
train <- sig_data[train_samples,]
test <- sig_data[-train_samples,]
model2 <- lm(Agr_area~.,data=train)
```

Summary:
```{r}
summary(model2)
```
Performance:
```{r}
sigma(model2)
```

```{r}
confint(model2)
```

PLots:
```{r}
plot(model2, 1)
```

```{r}
par(mar=c(2,2,2,2))
hist(model2$residuals)
```

```{r}
qqnorm(model2$residuals,ylab = "Residuals")
qqline(model2$residuals)
```

Test set prediction accuracy:
```{r}
pred <- model2 %>%
  predict(test)
R2 <- R2(pred,test$Agr_area)
R2
```

```{r}
RMSE <- RMSE(pred,test$Agr_area)
RMSE
```
## Data visualisation:

Data visualisation:
```{r}
densityplot(data$Agr_area)
```

Kurtosis and Skewness
```{r}
s = skewness(data$Agr_area)
k = kurtosis(data$Agr_area)
s
k
#K > 0 thus, verfified
```

Visuaisation:
```{r}
ggplot(data,aes(x=Agr_area,y=meadows,color=Arable))+geom_point()+ggtitle("Agr_area vs meadows, Color = Arable")
```


Quartiles:
```{r}
IQR(data$Agr_area)
percentile_i = quantile(data$Agr_area, c(.25,.50,.75))
percentile_i
```

Boxplots:
```{r}
boxplot(data$Agr_area, main="Agr_area boxplot", ylab = "Agr_area")
```

Histogram:
```{r}
hist(data$Agr_area, main = "Agr_area", xlab = "Agr_area", ylab = "frequency")
```

Pair plots:
```{r}
pairs(data[,])
```

Multi-box plot
```{r}
boxplot(data[, 2:6], main="Multiple-boxplot for Agricultural area dataset", ylab = "units")
```

bar plot:
```{r}
df1 <- data %>%
  filter(part_area >= 20)
ggplot(df1,aes(x=Country))+geom_bar()+ggtitle("Countries with over 1/5th agricultural area")
```

Scatter plot:
```{r}
ggplot(data, aes(x=meadows,y=Agr_area))+geom_point()
```

```{r}
ggplot(data, aes(x=crops,y=Agr_area))+geom_point()
```

```{r}

```

```{r}

```