---
title: "Lab exercise 2"
author: "Makesh Srinivasan"
date: "8/10/2021"
output: html_document
---
Registration number: 19BCE1717 <br />
Faculty: Dr. C. Sweetlin Hemalatha <br />
Slot: L39 + L40 <br />
Course code: CSE3505 <br />

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Ex 2A) Data frames in R
### Creating a data frame
1. Create a data frame by considering 5 students and 4 events. Each event has a maximum score of 10. If a student participates in an event, its entry contains the score value and 0 otherwise.
```{r}
students <- c("Makesh", "Anthra", "Bettina", "Chandhru", "Shyam")
event_1 <- c(10,10,6,8,10)
event_2 <- c(5,7,9,0,0)
event_3 <- c(7,9,0,8,6)
event_4 <- c(9,7,8,10,10)
q1a_dataframe <- data.frame(students, event_1, event_2, event_3, event_4)
q1a_dataframe
```

2.	View the contents of the data frame.
```{r}
# View(q1a_dataframe) should be used to view the data frame but since it is not being displayed on the HTML page I printed normally
q1a_dataframe
```

3.	Find the total score of each participant.
```{r}
total_score <- rowSums(q1a_dataframe[2:5])
total_score
```

4.	Append a column to include the total score of the participants and view the data frame.
```{r}
q1a_dataframe["total_score"] <- total_score
q1a_dataframe
```

5.	Find the maximum score and display the name of the participant who scored it.
```{r}
print(paste("Maximum score = ", max(q1a_dataframe$total_score)))
print(paste("Name = ", q1a_dataframe$students[which.max(q1a_dataframe$total_score)]))
```

6.	Compute the average score of each event and append it as a new row in the data frame.
```{r}
q1a_dataframe <- rbind(q1a_dataframe, data.frame(students="Average_score", event_1=mean(q1a_dataframe$event_1), event_2=mean(q1a_dataframe$event_2), event_3=mean(q1a_dataframe$event_3), event_4=mean(q1a_dataframe$event_4), total_score=mean(q1a_dataframe$total_score)))
q1a_dataframe
```

7.	Store the details in a comma separated values (csv) file. Also suppress the row numbers.
```{r}
write.csv(q1a_dataframe, "q1a_dataframe.csv", row.names = FALSE)
# I have manually verified that the CSV file has been created in the current working directory
```

### Indexing and Slicing data frames

8.	Read the content of ‘Events.csv’ in a data frame and view it.
```{r}
# My file was not named as "Events.csv". It was named as "q1a_dataframe.csv"
df2 <- read.csv("q1a_dataframe.csv")
# View(df2) should be used to view the data frame but since it is not being displayed on the HTML page I printed normally
df2
```

9.	Access the scores of participants in event2 using the column name.
```{r}
df2["event_2"]
```

10.	Use index number to retrieve the same data.
```{r}
df2[3]
```

11.	Extract the score of third participant in event3.
```{r}
df2[3,4]
```

12.	Extract the scores of the first and second participant in all the events.
```{r}
df2[c(1:2), c(2:5)]
```

13.	Display the names and total scores of all participants.
```{r}
df2[c(1:5), c(1,6)]
```

14.	Make the column “name” as the row index of the data frame.
```{r}
# student column is the 'name' as asked in the question
row.names(df2) <- df2$students
print(row.names(df2))
```
```{r}
df2
```

15.	Display the names of the students participated in event3.
```{r}
boolv <- df2["event_3"]>0
row.names(df2)[boolv]
```

16.	Obtain the names whose total score is above its average.
```{r}
boolv2 <- df2["total_score"]>df2[6,5]
row.names(df2)[boolv2]
```

## Ex 2B) Basic Operations in data frame
### MASS package contains a data frame called ‘survey’ which contains the responses of 237 Statistics I students at the University of Adelaide to a number of questions. Solve the following 10 questions.

1.	Install the package MASS.
```{r}
# installed using install.packages("MASS")
library(MASS)
```

2.	Import the package MASS.
```{r}
# stored survey dataframe in df3
library(MASS)
data(survey)
df3 <- survey
```

3.	Display the structure of the data survey.
```{r}
# str(df3) can be used too but to make it clear and maintain consistency in the answer, I have used str(survey) for this an the following questions
str(survey)
```

4.	Check the class and type of the data set survey in MASS.
```{r}
print(paste("Class = ", class(survey)))
print(paste("Type = ", typeof(survey)))
```

5.	Get the number of rows and columns of the survey data frame.
```{r}
print(paste("Rows = ", nrow(survey)))
print(paste("Columns = ", ncol(survey)))

```

6.	Get the dimension of the survey data frame.
```{r}
dim(survey)
```

7.	Provide the statistical summary of the data frame.
```{r}
summary(survey)
```

8.	Display the column names of the survey data frame
```{r}
names(survey)
```

9.	Retrieve the top 3 rows from the data frame.
```{r}
survey[1:3,]
```

10.	Extract the bottom 2 rows from the data frame.
```{r}
survey[(dim(survey)[1]-1):(dim(survey)[1]),]
```






