---
title: "Lab 8"
author: "Makesh Srinivasan"
date: "10/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Regression analysis
Import the dataset "Forestfires.csv"
```{r}
data <- read.csv("forestfires.csv")
head(data)
```
#### Process the dataset before building a linear regression model
1) Replace non-numeric values with numeric data
1.1) The day column is replaced
```{r}
data[data$day=="mon", 'day'] <- as.numeric(1)
data[data$day=="tue", 'day'] <- as.numeric(2)
data[data$day=="wed", 'day'] <- as.numeric(3)
data[data$day=="thu", 'day'] <- as.numeric(4)
data[data$day=="fri", 'day'] <- as.numeric(5)
data[data$day=="sat", 'day'] <- as.numeric(6)
data[data$day=="sun", 'day'] <- as.numeric(7)
head(data, 7)
```
1.2) The month column is replaced
```{r}
data[data$month=="jan", 'month'] <- as.numeric(1)
data[data$month=="feb", 'month'] <- as.numeric(2)
data[data$month=="mar", 'month'] <- as.numeric(3)
data[data$month=="apr", 'month'] <- as.numeric(4)
data[data$month=="may", 'month'] <- as.numeric(5)
data[data$month=="jun", 'month'] <- as.numeric(6)
data[data$month=="jul", 'month'] <- as.numeric(7)
data[data$month=="aug", 'month'] <- as.numeric(8)
data[data$month=="sep", 'month'] <- as.numeric(9)
data[data$month=="oct", 'month'] <- as.numeric(10)
data[data$month=="nov", 'month'] <- as.numeric(11)
data[data$month=="dec", 'month'] <- as.numeric(12)
head(data, 7)
```
The dataset is now entirely numeric:
```{r}
data <- as.data.frame(apply(data, 2, as.numeric))
str(data)
```
2) Check for null or NaN values
```{r}
print(paste("Number of missing values = ", sum(is.na(data))))
```
Since there are no null or NaN values, the dataset is clean. We can now start plotting and modelling

#### Plot
Pair-wise plot with each column and the label (area)
```{r}
library(ggplot2)
par(mar=c(2,2,2,2))
ggplot(data,aes(x=X,y=area))+geom_point()
ggplot(data,aes(x=Y,y=area))+geom_point()
ggplot(data,aes(x=month,y=area))+geom_point()
ggplot(data,aes(x=day,y=area))+geom_point()
ggplot(data,aes(x=FFMC,y=area))+geom_point()
ggplot(data,aes(x=DMC,y=area))+geom_point()
ggplot(data,aes(x=DC,y=area))+geom_point()
ggplot(data,aes(x=ISI,y=area))+geom_point()
ggplot(data,aes(x=temp,y=area))+geom_point()
ggplot(data,aes(x=RH,y=area))+geom_point()
ggplot(data,aes(x=wind,y=area))+geom_point()
ggplot(data,aes(x=rain,y=area))+geom_point()
```
Correlation between the features and the label column - area
```{r}
cor(data$X,data$area)
cor(data$Y,data$area)
cor(data$month,data$area)
cor(data$day,data$area)
cor(data$FFMC,data$area)
cor(data$DMC,data$area)
cor(data$DC,data$area)
cor(data$ISI,data$area)
cor(data$temp,data$area)
cor(data$RH,data$area)
cor(data$wind,data$area)
cor(data$rain,data$area)
```

Model with all columns
```{r}
model <- lm(area~.,data=data)
model
```

```{r}
summary(model)
```

```{r}
sigma(model)
```

```{r}
confint(model)
```

```{r}
#########Diagnostics##########
#Evaluating the residuals
plot(model,1)

model1 <- lm(log(area)~month,data=data)
plot(model1,1)

```

```{r}
install.packages('tidyverse')
install.packages('caret')
library(tidyverse)
library(caret)
```

```{r}
set.seed(123)
train_samples <- Boston$medv %>%
  createDataPartition(p=0.8,list=FALSE)
head(train_samples)
train <- Boston[train_samples,]
test <- Boston[-train_samples,]
```

```{r}
```

```{r}
```

```{r}
```
