---
title: "fat notes"
author: "Makesh Srinivasan"
date: "12/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(ggplot2)
library(tidyverse)
library(MASS)
library(dplyr)
library(tidyr)
library(stringr)
library(caret)
library(lattice)
library(e1071)
```

```{r}
library(MASS)
library(dplyr)
survey <- survey
df <- unique(survey)
df <- na.omit(survey)
library(ggplot2)
library(tidyverse)
View(df)
```

```{r}
ggplot(data=df, aes(x=Sex)) + geom_bar()+ggtitle("Male and Female participants")
```

```{r}
ggplot(data=df, aes(x=Wr.Hnd, y=NW.Hnd)) + geom_point() + geom_line() + facet_wrap(~Sex)+geom_smooth(method=lm)
```

```{r}
ggplot(data=df, aes(x=Sex, y=Wr.Hnd))+geom_boxplot()+geom_point(aes(size=Age, colour=Wr.Hnd)) + labs(title = "title")
# use continuous for size in geom_point(aes()) and colour is discrete/continuous doesn't matter
```

```{r}
ggplot(df, aes(x="", y=age,fill=val))+geom_bar(width=1,stat ='identity')+coord_polar("y",start=0)+ggtitle("Smoking habits of male left handers based on 'Age'")
```

```{r}
age <- factor(df$Age)
val <- factor(df$Smoke)
typeof(val)
```


```{r}
ggplot(df, aes(x=Fold))+geom_bar(aes(fill=Fold))+labs(title="something")+facet_wrap(~Sex)
```

```{r}
ggplot(df ,aes(x=Wr.Hnd,y=Pulse))+geom_point(aes(size=Height,color=Sex))

```

Plot the distribution of height of the students with filled colour based on gender with facet based on left and write handers.
```{r}
ggplot(df,aes(x=Height, fill=Sex))+geom_histogram(binwidth = 1)+labs(title = "Distribution of height",y='Frequency')+facet_wrap(~W.Hnd)
```

```{r}
gf = factor(df$W.Hnd,levels = c("Left","Right"),labels = c("Left","Right"))
gf

```


Draw the distribution of smoking habits of male left handers based on age using pie chart.

```{r}
df%>%
  filter(Sex == "Male" & W.Hnd == "Left") %>%
  ggplot(aes(x="", y=Age, fill=Smoke))+geom_bar(width=1, stat="identity")+coord_polar("y", start=0)+ggtitle("Smoking habits of male left handers based on 'Age'")+ theme(axis.text = element_blank(),axis.ticks=element_blank(),panel.grid = element_blank())



#age <- factor(df2$Age)
#val <- factor(df2$Smoke)

#ggplot(df2, aes(x="",y=age,fill=val))+geom_bar(width=1,stat ='identity')+coord_polar("y",start=0)+ggtitle("Smoking habits of male left handers based on 'Age'")+ theme(axis.text = element_blank(),axis.ticks=element_blank(),panel.grid = element_blank())
```

```{r}
df2 <- survey %>%
  filter(Sex == "Male" & W.Hnd == "Left")

age <- factor(df2$Age)
val <- factor(df2$Smoke)

ggplot(df2, aes(x="",y=age,fill=val))+geom_bar(width=1,stat ='identity')+coord_polar("y",start=0)+ggtitle("Smoking habits of male left handers based on 'Age'")+ theme(axis.text = element_blank(),axis.ticks=element_blank(),panel.grid = element_blank())
```

```{r}
survey <- survey
newsurvey <- unique(survey)
newsurvey <- na.omit(survey)
```

```{r}
gender = table(newsurvey$Sex)
gender
```


Find the relative frequency distribution of male and female participants and display them with the precision of two decimal places.
```{r}
old=options(digits = 2)
sex_freq = table(newsurvey$Sex)
samplesize = nrow(newsurvey)
sex_relfreq = sex_freq/samplesize
cbind(sex_relfreq)
```

```{r}
newsurvey%>%
  group_by(Sex) %>%
  summarise(freq=n()) %>%
  mutate(rel.freq = freq/sum(freq))
```

```{r}
newsurvey%>%
  filter(W.Hnd=="Right") %>%
  group_by(Sex) %>%
  summarise(freq=n()) 
  #mutate(rel.freq = freq/sum(freq))
```

Find the range of students’ age participated in the survey.
Break the age range into non-overlapping sub-intervals by defining a sequence of equal distance break points of 10 by rounding the range to nearest integer.
Find the distribution of the age range according to the sub-intervals with cut with its right boundary opened. Display it in column form.
```{r}
range(newsurvey$Age)

breaks=seq(10,80,by=10)
breaks

sub_intervals = cut(newsurvey$Age, breaks, right=FALSE)
sub_intervals <- table(sub_intervals)
cbind(sub_intervals)
```

Which age range of students has mostly participated in the survey.
```{r}
print(paste("Max = ", max(sub_intervals)))
which.max(sub_intervals)
```

```{r}
percentile_age=quantile(newsurvey$Age, c(.20,.50,.70))
percentile_age
```


Which category of exercise students has the maximum writing hand span?
```{r}
max_hnd <- newsurvey%>%
  group_by(Exer)%>%
  summarise(max = max(Wr.Hnd))
max_hnd
newsurvey[which.max(newsurvey$Wr.Hnd),]$Exer
```

```{r}
cols = colnames(data)
for (c in cols) {
  print(paste("Column", c, ": ", cor(data[[c]],data$price))) 
}
```

```{r}
densityplot(data$V1)
```

```{r}
densityplot(df$Height)
```

```{r}
print(paste("Kurtosis = ", kurtosis(df$Height)))
print(paste("Skewness = ", skewness(df$Height)))
```

```{r}
ggplot(df, aes(Height)) + geom_boxplot()
```

```{r}
cor(data)
```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```


### Predicting the Whole.weight from the Length of the abalone

```{r}
data <-  subset(data,select = -c(Sex))
colnames(data)
ggplot(abalone,aes(x=Length,y=Whole.weight))+geom_point()
ggplot(abalone,aes(x=Length,y=Whole.weight))+geom_point()+geom_smooth()
ggplot(abalone,aes(x=Length,y=Whole.weight))+geom_point()+geom_smooth(method='lm',se=FALSE)

```

### Correlation between weight and length of the abalone

```{r}
cor(abalone$Length,abalone$Whole.weight)
```

### Train test split

```{r}
library(tidyverse)
library(caret)

set.seed(123)
train_ <- abalone$Whole.weight %>%
  createDataPartition(p=0.8,list=FALSE)

train <- abalone[train_,]
test <- abalone[-train_,]
```

### Building a linear model

```{r}
model = lm(Length~Whole.weight,data=train)
model
```

```{r}
summary(model)
```
### Confidence interval of the model

```{r}
confint(model)
```

### Confidence interval

```{r}
Whole.weight <- 000
test <- data.frame(Whole.weight)
confid <- predict(model,test,level = .95,interval = "confidence")
confid
```

### Prediction interval

```{r}
predid <- predict(model,test,level = .95,interval = "prediction")
predid
```
#### In residuals Median is very close to 0 and thus mean is 0. The t factor has a significance of 0.001 and thus it is statistically significant relationship

#### F statistic's value is also statistically significant

### Prediction
```{r}
pred <- model %>%
  predict(test)
```

### Diagnostics
```{r}
plot(abalone$Length,abalone$Whole.weight)
```

### Evaluating the residuals
```{r}
plot(model,1)
```

### Errors

```{r}
val = sigma(model)
perc = val*100/mean(abalone$Whole.weight)
sprintf("The Residual Standard Deviation is %f and its percentage is %f%%",val,perc)
```


# Linear Regression in multiple variables

X = entire dataset except Length
Y = Length

### Finding correlation between the variables

```{r}
cor(abalone)
```
### From the correlation matrix -> determining to find length using the other variables

(For next model -> Drop height and rings column as they have low correlation)

```{r}
pairs(abalone)
#Pairplot to see the relation between the various attributes. 
```

```{r}
print("Diameter vs Length")
ggplot(abalone,aes(x=Diameter,y=Length))+geom_point()+geom_smooth(method='lm',se=FALSE)
```

```{r}
print("Whole weight vs Length")
ggplot(abalone,aes(x=Whole.weight,y=Length))+geom_point()+geom_smooth(method='lm',se=FALSE)
```

```{r}
print("Shucked weight vs Length")
ggplot(abalone,aes(x=Shucked.weight,y=Length))+geom_point()+geom_smooth(method='lm',se=FALSE)
```

```{r}
print("Viscera Weight vs Length")
ggplot(abalone,aes(x=Viscera.weight,y=Length))+geom_point()+geom_smooth(method='lm',se=FALSE)
```
```{r}
print("Shell Weight vs Length")
ggplot(abalone,aes(x=Shell.weight,y=Length))+geom_point()+geom_smooth(method='lm',se=FALSE)
```

```{r}
print("Length vs Rings")
ggplot(abalone,aes(x=Rings,y=Length))+geom_point()+geom_smooth(method='lm',se=FALSE)
```

```{r}
print("Height vs Length")
ggplot(abalone,aes(x=Height,y=Length))+geom_point()+geom_smooth(method='lm',se=FALSE)
```

### Splitting into train and test

```{r}
set.seed(123)
train_ <- abalone$Length %>%
  createDataPartition(p=0.8,list=FALSE)

train <- abalone[train_,]
test <- abalone[-train_,]
```

### Building the model

```{r}
model <- lm(Length~.,data=train)
summary(model)
```

### Making prediction

```{r}
pred <- model %>%
  predict(test)
```

### Model Performance 
```{r}
#model performance
RMSE <- RMSE(pred,test$Length)
RMSE
R2 <- R2(pred,test$Length)
R2
```

### Inference 
The t statistic is large for 3 variables and small for the rest, the attributes for which it is large will contribute constructively for model building. 

Residual Standard Error : 0.01877 on 3335 degrees of freedom
Multiple R-squared:  0.9756
Adjusted R-squared:  0.9755
F-statistic: 1.901e+04 on 7 and 3335 DF
p-value: < 2.2e-16

Median value is close to 0 and thus this means that mean will be close to 0, making the mean of the residuals almost 0. 

### Residual Standard Deviation

```{r}
val = sigma(model)
perc = sigma(model)*100/mean(train$Length)
sprintf("The Residual Standard Deviation is %f and its percentage is %f%%",val,perc)
```

### Confidence interval

```{r}
confint(model)
```
```{r}
par(mar=c(2,2,2,2))
hist(model$residuals)
```
### Normal qq plot of residuals

```{r}
qqnorm(model$residuals,ylab = "Residuals")
qqline(model$residuals)
```

# Building Model with statistically significant columns

### Building another model after dropping columns with no significance = Height, Whole weight, Shell Weight, Rings

```{r}
abalone1 <- subset(abalone,select=-c(Height,Rings,Shell.weight,Whole.weight))
```

### Splitting into train and testing sets

```{r}
set.seed(123)
train_ <- abalone1$Length %>%
  createDataPartition(p=0.8,list=FALSE)

train1 <- abalone1[train_,]
test1 <- abalone1[-train_,]
```


### Building the model
```{r}
model1 <- lm(train1$Length~.,data=train1)
```

```{r}
summary(model1)
```
### Final Metrics

```{r}
RMSE <- RMSE(pred,test1$Length)
R2 <- R2(pred,test1$Length)
sprintf("The RMSE error is %f and the R2 error is %f",RMSE,R2)
```

