---
title: "Lab 4: Dataset cleaning"
author: "Makesh Srinivasan"
date: "8/24/2021"
output: html_document
---
Registration number: 19BCE1717 <br>
Faculty: Dr. C. Sweetlin Hemalatha <br>
Slot: L39 + L40 <br>
Course code: CSE3505 <br>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Dataset cleaning
#### The dataset used for this lab exercise is https://www.kaggle.com/uciml/iris 
#### The dataset is clean as it is and therefore, I introduced NA values randomly, changed datatypes and duplicated entries to perform dataset cleaning on this. There are 6 columns and 150 entries for 3 different species of iris
```{r}
# import the libraries
library(tidyverse)
library(dplyr)
```

```{r}
# import the dataset
original_iris <- read.csv("Iris.csv")
summary(original_iris)
```

### The dataset is clean, i.e, there are no missing values or any duplicate entries. Thus, they will be added in the following sections
```{r}
print(paste("Number of missing values = ", sum(is.na(original_iris))))
print(paste("Number of duplicate entries = ", sum(table(original_iris$Id)-1)))
```

### Adding Missing values (NA)
```{r}
# inject NA values
dirty_iris <- original_iris %>%
  group_by(Species) %>%
  mutate(SepalLengthCm = replace(SepalLengthCm, sample(row_number(),size = ceiling(0.2 * n()), replace = TRUE), NA))%>%
  mutate(SepalWidthCm = replace(SepalWidthCm, sample(row_number(),size = ceiling(0.2 * n()), replace = TRUE), NA))%>%
  mutate(PetalLengthCm = replace(PetalLengthCm, sample(row_number(),size = ceiling(0.2 * n()), replace = TRUE), NA))%>%
  mutate(PetalWidthCm = replace(PetalWidthCm, sample(row_number(),size = ceiling(0.2 * n()), replace = TRUE), NA))

print(paste("Number of missing values = ", sum(is.na(dirty_iris))))
print(paste("Number of duplicate entries = ", sum(table(dirty_iris$Id)-1)))
```

### Changing the datatype of a column from num to chr
```{r}
# Datatype change from num to chr for column SepalLengthCm
dirty_iris <- dirty_iris %>% 
  mutate(SepalLengthCm = as.character(SepalLengthCm))
str(dirty_iris)
```

### Adding duplicate entries to the dataset
```{r}
duprow = dirty_iris[1,]
for(i in 1:25){
  dirty_iris = rbind(dirty_iris, duprow)
}

print(paste("Number of missing values = ", sum(is.na(dirty_iris))))
print(paste("Number of duplicate entries = ", sum(table(dirty_iris$Id)-1)))
```

```{r}
# the dimension is now 175 with 6 columns - an increase from 150 rows due to duplicates
dim(dirty_iris)
```
### The dataset is now entirely dirty! the following sections are dataset cleaning
#### Drop the duplicate entries
```{r}
clean_iris <- unique(dirty_iris)
print(paste("Number of rows after removing duplicates = ", nrow(clean_iris)))
```

#### The numbers are in character datatype and to be processed they need to be converted to numeric datatype
```{r}
print("Before changing the datatype")
str(clean_iris)
clean_iris <- clean_iris %>% 
  mutate(SepalLengthCm = as.numeric(SepalLengthCm))
print("After changing the datatype")
str(clean_iris)
```

#### There are many NA values in the column SepalLengthCm that needs to be filled
```{r}
unique(clean_iris$SepalLengthCm)
```
#### Find mean, median, min and max
```{r}
print(paste("Mean value of the column SepalLengthCm = ", mean(clean_iris$SepalLengthCm,na.rm = TRUE)))
print(paste("Median value of SepalLengthCm = ", median(clean_iris$SepalLengthCm,na.rm = TRUE)))
print(paste("Maximum value of SepalLengthCm = ", max(clean_iris$SepalLengthCm,na.rm = TRUE)))
print(paste("Minimum value of SepalLengthCm = ", min(clean_iris$SepalLengthCm,na.rm = TRUE)))
```

#### The mean and median values are very close while min and max values are having a wider difference in magnitude. Hence, the use of mean or median in substituting for a NA value in this column is a good practice
```{r}
# Used meadian to fill the missing values
clean_iris <- clean_iris %>%
  mutate(SepalLengthCm = replace(SepalLengthCm, is.na(SepalLengthCm), median(SepalLengthCm, na.rm = TRUE)))
```
#### Now verify that there are no missing or NA values in this column (SepalLengthCm)
```{r}
print(paste("The number of missing values (column SepalLengthCm) are = ", sum(is.na(clean_iris$SepalLengthCm))))
```

#### We repeat this for the next column - SepalWidthCm
#### There are some missing values in this column
```{r}
unique(clean_iris$SepalWidthCm)
```
#### The datatype is in numeric format so we can directly move to filling missing values
```{r}
#### Find mean, median, min and max
print(paste("Mean value of the column SepalWidthCm = ", mean(clean_iris$SepalWidthCm,na.rm = TRUE)))
print(paste("Median value of SepalWidthCm = ", median(clean_iris$SepalWidthCm,na.rm = TRUE)))
print(paste("Maximum value of SepalWidthCm = ", max(clean_iris$SepalWidthCm,na.rm = TRUE)))
print(paste("Minimum value of SepalWidthCm = ", min(clean_iris$SepalWidthCm,na.rm = TRUE)))
```

#### Again, the mean and median are very close compared to the min or max values of this column. Hence, either mean or median can be used to replace missing values
```{r}
# used median value to replace the missing values
clean_iris <- clean_iris %>%
  mutate(SepalWidthCm = replace(SepalWidthCm, is.na(SepalWidthCm), median(SepalWidthCm, na.rm = TRUE)))
```
#### Now verify that there are no missing or NA values in this column (SepalWidthCm)
```{r}
print(paste("The number of missing values (column SepalWidthCm) are = ", sum(is.na(clean_iris$SepalWidthCm))))
```



#### We repeat this for the next column - PetalLengthCm
#### There are some missing values in this column
```{r}
unique(clean_iris$PetalLengthCm)
```
#### The datatype is in numeric format so we can directly move to filling missing values
```{r}
#### Find mean, median, min and max
print(paste("Mean value of the column PetalLengthCm = ", mean(clean_iris$PetalLengthCm,na.rm = TRUE)))
print(paste("Median value of PetalLengthCm = ", median(clean_iris$PetalLengthCm,na.rm = TRUE)))
print(paste("Maximum value of PetalLengthCm = ", max(clean_iris$PetalLengthCm,na.rm = TRUE)))
print(paste("Minimum value of PetalLengthCm = ", min(clean_iris$PetalLengthCm,na.rm = TRUE)))
```

#### Again, the mean and median are quite close compared to the min or max values of this column. Hence, either mean or median can be used to replace missing values
```{r}
# used mean value to replace the missing values
clean_iris <- clean_iris %>%
  mutate(PetalLengthCm = replace(PetalLengthCm, is.na(PetalLengthCm), mean(PetalLengthCm, na.rm = TRUE)))
```
#### Now verify that there are no missing or NA values in this column (PetalLengthCm)
```{r}
print(paste("The number of missing values (column PetalLengthCm) are = ", sum(is.na(clean_iris$PetalLengthCm))))
```



#### Only one column is left. We repeat this for the last column - PetalWidthCm
#### There are some missing values in this column
```{r}
unique(clean_iris$PetalWidthCm)
```
#### The datatype is in numeric format so we can directly move to filling missing values
```{r}
#### Find mean, median, min and max
print(paste("Mean value of the column PetalWidthCm = ", mean(clean_iris$PetalWidthCm,na.rm = TRUE)))
print(paste("Median value of PetalWidthCm = ", median(clean_iris$PetalWidthCm,na.rm = TRUE)))
print(paste("Maximum value of PetalWidthCm = ", max(clean_iris$PetalWidthCm,na.rm = TRUE)))
print(paste("Minimum value of PetalWidthCm = ", min(clean_iris$PetalWidthCm,na.rm = TRUE)))
```

#### Again, the mean and median are quite close compared to the min or max values of this column. Hence, either mean or median can be used to replace missing values
```{r}
# used mean value to replace the missing values
clean_iris <- clean_iris %>%
  mutate(PetalWidthCm = replace(PetalWidthCm, is.na(PetalWidthCm), mean(PetalWidthCm, na.rm = TRUE)))
```
#### Now verify that there are no missing or NA values in this column (PetalWidthCm)
```{r}
print(paste("The number of missing values (column PetalWidthCm) are = ", sum(is.na(clean_iris$PetalWidthCm))))
```

### The dataset is now completely clean!
```{r}
# verify that the dataset is clean by checking if there are any NA values
any(is.na(clean_iris))
```
```{r}
print(paste("Number of missing values after cleaning = ", sum(is.na(clean_iris))))
print(paste("Number of duplicate entries after cleaning = ", sum(table(clean_iris$Id)-1)))
```

### The dataset can be used for the next phase such as to visualise the different species
```{r}
summary(clean_iris)
```
```{r}
print("The different species present are: ")
df <- clean_iris%>%
  group_by(Species) %>%
  select(Species)
unique(df)
```